TODO

### 2024-12-13 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
draft stats and figures

form <- function(model) {
  dep <- as.character(model$formula[2])
  ind <- as.character(model$formula[3])

  str_c(dep, " ~ ", ind) %>%
    str_replace_all(" ", "") %>%
    str_replace_all("\\*", " $\\\\times$ ") %>%
    str_replace_all("\\+", "&nbsp;+ ") %>%
    str_replace_all("~", "$\\\\sim$ ")
}

@@

@@

* save model summaries

@@

@@

@@

```{r}
 clean_1 |> 
  ggplot(aes(x = trial+1, y = as.numeric(correct), fill=type)) +
  stat_summary(fun.data="mean_cl_boot", geom="bar") +
    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
  scale_fill_manual(values=time_palette)+
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

@@

@@

```{r}
 clean_1 |> 
  mutate(time=ifelse(is.na(time),180000,time)) %>% #treat the one that timed out as being max time
  ggplot(aes(x = trial+1, y = time/1000, color=type)) +
  geom_jitter(width=.2, height=0, alpha=.5)+
  #geom_line(aes(group=game))+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Time to answer in seconds")+
  coord_cartesian(ylim=c(0,60))+
  #ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
```

@@

clean_1 |> 
  ggplot(aes(x = trial+1, y = words, color=type)) +
    geom_jitter(width=.2, height=0, alpha=.5)+
  #geom_line(aes(group=game))+
 # facet_wrap(.~correct) + 
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  #coord_cartesian(ylim=c(0,60))+
  #ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

@@

@@

```{r}
sims_1 |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |>
  ggplot(aes(x = source, y = sim, color = target1)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
  theme(legend.position = "bottom")

@@

```{r}
sims_1 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 3) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |> 
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r}
sims_1 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r}
sims_1 |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```
# Experiment 1 Discussion

@@

```{r}
clean_2 |>
  ggplot(aes(x = trial + 1, y = correct)) +
  stat_summary(aes(fill = type), fun.data = "mean_cl_boot", geom = "bar") +
  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_fill_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(data = clean_2 |> filter(trial > 3), method = "lm", color = "black") +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1, .1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  theme(
    aspect.ratio = 4 / 5,
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 9, color = "black"),
    legend.position = "none"
  )
```

@@

@@

```{r}
clean_2 |> ggplot( aes(x = trial + 1, y = time / 1000, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = clean_2 |> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Time (seconds)") +
  theme(legend.position = "none")
```

@@

clean_2 |> ggplot( aes(x = trial + 1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = clean_2|> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  theme(legend.position = "none")
```

@@

```{r}
sims_2 |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |>
  ggplot(aes(x = source, y = sim, color = target1)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
  theme(legend.position = "bottom")

```{r}
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r}
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r}
sims_2 |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

>>>>

@@

```{r}
stats <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
}

stats_text <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 2], "  ", model[row, 3])
}

form <- function(model_form) {
  dep <- as.character(model_form$formula[2])
  ind <- as.character(model_form$formula[3])

  str_c(dep, " ~ ", ind) |>
    str_replace_all(" ", "") |>
    str_replace_all("\\*", " $\\\\times$ ") |>
    str_replace_all("\\+", "&nbsp;+ ") |>
    str_replace_all("~", "$\\\\sim$ ")
}

@@

@@

```{r interface, fig.env = "figure*", fig.pos = "t!", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TODO \\label{game}", cache=FALSE}

@@

## Statistical analysis

Statistical analyses were run in brms with weakly informative priors. We present the estimate and 95% credible intervals.

@@

```{r accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
clean_2 |>
    filter(type!="practice") |> mutate(trial=trial-3, expt="Expt 2") |> 
  bind_rows(clean_1 |> filter(type!="practice") |> mutate(trial=trial-1, expt="Expt 1")) |> 
  ggplot(aes(x = trial, y = correct)) +
  stat_summary(aes(fill = type), fun.data = "mean_cl_boot", geom = "bar") +
  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_fill_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(method = "lm", color = "black") +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  xlab("Trial") +
  ylab("Accuracy") +
  scale_x_continuous(breaks = seq(1, 16, 4), expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  facet_grid(expt~.)+
  theme(legend.position = "none",
        panel.grid = element_blank(),
        strip.background = element_blank())
```

@@

```{r}
acc_mod_1 <- read_rds(here(mod_results, "acc_1.rds")) |> mutate(across(c(`Estimate`,`lower`,`upper`), ~exp(.)))

acc_form <- read_rds(here(mod_form,"acc_1.rds"))
```

We fit a mixed effects model of accuracy (`r acc_form |> form()`). Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`) and accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`).

@@

```{r}
speed_mod_1 <- read_rds(here(mod_results, "speed_1.rds"))

speed_form <- read_rds(here(mod_form, "speed_1.rds"))
```

We ran a Bayesian mixed effects model of how long each trial took over time: `r speed_form |> form()`. The first trial was `r stats_text(speed_mod_1, 1)`, and children got faster over time (`r stats_text(speed_mod_1, 2)`).

@@

```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }

clean_2 |>
    filter(type!="practice") |> mutate(trial=trial-3, expt="Expt 2") |> 
  bind_rows(clean_1 |> filter(type!="practice") |> mutate(trial=trial-1, expt="Expt 1")) |> ggplot( aes(x = trial, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  geom_smooth(method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(expt~.)+
  theme(legend.position = "none",
        panel.grid = element_blank(),
        strip.background = element_blank())
```

```{r}
description_mod_1 <- read_rds(here(mod_results, "description_1.rds"))

description_form <- read_rds(here(mod_form, "description_1.rds"))

@@

We ran a Bayesian mixed effects model of how long of a description the "teller" produced: `r description_form |> form()`. The initial length was `r stats_text(description_mod_1, 1)` and description length was relatively stable over time (`r stats_text(description_mod_1, 2)`).

@@

```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }

sims_2 |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |> mutate(expt="Expt 2") |> 
  bind_rows(sims_1 |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |> mutate(expt="Expt 1")) |> 
  ggplot(aes(x = source, y = sim, color = target1)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
  facet_grid(expt~.)+
  theme(legend.position = "bottom",
        panel.grid = element_blank(),
        strip.background = element_blank())

@@

```{r, include=F}
sims_1 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 3) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |> 
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r include=F}
sims_1 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")

```{r, include=F}
sims_1 |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

There were too few trials to really look at change over time in this short experiment. However, we were able to look at ...

We tried various ways of looking at the semantic similarities of descriptions children produced. TODO talk about SBERT

# Experiment 1 Discussion

@@

@@

```{r}
acc_mod_2 <- read_rds(here(mod_results, "acc_2.rds")) |> mutate(across(c(`Estimate`,`lower`,`upper`), ~exp(.)))

 Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`).

@@

```{r}
speed_mod_2 <- read_rds(here(mod_results, "speed_2.rds"))

 The first trial was `r stats_text(speed_mod_2, 1)`, and children got faster over time (`r stats_text(speed_mod_2, 2)`).

@@

```{r}
description_mod_2 <- read_rds(here(mod_results, "description_2.rds"))

 The initial length was `r stats_text(description_mod_2, 1)` and description length was relatively stable over time (`r stats_text(description_mod_2, 2)`).

@@

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "TODO \\label{accuracy}" }
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r, include=F}
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r, include=F}
sims_2 |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

```{r}
sim_across_between_2_results <- read_rds(here(mod_results, "sim_across_between_2.rds"))

sim_across_between_form<- read_rds(here(mod_form, "sim_across_between_2.rds"))

sim_across_2_results <- read_rds(here(mod_results, "sim_across_2.rds"))

sim_across_form<- read_rds(here(mod_form, "sim_across_2.rds"))

sim_to_next_2_results <- read_rds(here(mod_results, "sim_to_next_2.rds"))

sim_to_next_form<- read_rds(here(mod_form, "sim_to_next_2.rds"))

sim_to_last_2_results <- read_rds(here(mod_results, "sim_to_last_2.rds"))

sim_to_last_form<- read_rds(here(mod_form, "sim_to_last_2.rds"))
```

We tried various ways of looking at the semantic similarities of descriptions children produced. TODO talk about SBERT

As a coarse comparison, we repeated the analysis from Experiment 1, comparing the similarity of descriptions to the same target for the same-speaker, same-game, or different-game. We modeled this as `r sim_across_between_form |> form()`. Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person with the partnership (`r sim_across_between_2_results |> stats_text(3,3)`). The big differences in descriptions between games compared to within games is a measure of partner sensitivity -- children are more likely to use descriptions semantically similar to their partners than to another child.

With the greater length of experiment 2, we are also able to look for change over time, that could be indicative of convergence to shared descriptions within a game and divergence between games.

To look for convergence, we compare the utterances from the first three blocks to the descriptions in the last block: `r sim_to_last_form |> form()`. Over the first three blocks, descriptions do become increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`). Descriptions are more similar to the last block if they come from the same child who gaves the description in the last block (`r sim_to_last_2_results |> stats_text(3,3)`).

Another way to look for convergence is to look at the semantic distance between utterances in adjacent blocks: `r sim_to_next_form |> form()`. Although over time descriptions do get more similar to the last block utterance, the distance between adjacent block utterances is relatively constant `r sim_to_next_2_results |> stats_text(2,3)`.

Finally, we can see if descriptions in different games diverge across time: `r sim_across_form |> form()`. As the games progress, descriptions from different games become slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`).
